Demes: a standard format for demographic models

### 2022-05-27 paper.tex
Update software table

% local definitions
\newcommand{\ms}[0]{\texttt{ms}}
\newcommand{\msprime}[0]{\texttt{msprime}}
\newcommand{\stdpopsim}[0]{\texttt{stdpopsim}}
\newcommand{\demes}[0]{\texttt{demes}}
\newcommand{\demesdraw}[0]{\texttt{demesdraw}}
\newcommand{\Demes}[0]{\texttt{Demes}}
\newcommand{\moments}[0]{\texttt{moments}}
\newcommand{\dadi}[0]{\texttt{dadi}}
\newcommand{\fwdpy}[0]{\texttt{fwdpy11}}
\newcommand{\slim}[0]{\texttt{SLiM}}
\newcommand{\gadma}[0]{\texttt{GADMA}}
\newcommand{\tskit}[0]{\texttt{tskit}}

@@

The ever-increasing amount of genetic sequencing data from genetically and
geographically diverse species and populations has allowed us to infer complex
demography and study life history at fine scales.
An integral component to such population genetics studies is simulation.
Software to either simulate whole genome
sequences~\citep{thornton2014cpp,thornton2019-nu,staab2015scrm,
baumdicker2021-iu,kelleher2016efficient,haller2019slim}
or informative summary statistics of
diversity~\citep{gutenkunst2009inferring,kamm2017efficient,jouganous2017inferring}
have enabled the increasing complexity of genomic studies, with several software
packages capable handling large sample sizes, many interacting populations, and
deviations from panmictic random-mating assumptions.
This ability to infer and simulate such complex demographic scenarios, however,
has highlighted a major shortcoming in community standards:
the fragmented landscape of different ways to describe demographic
models makes it difficult to compare inferences made by different methods
and to reliably simulate from previously inferred models.
Inference results are typically reported in publications
via a combination of visual depiction,
a list of key parameters in tabular form and a discussion within the text.
Unfortunately these descriptions are often ambiguous, and
implementing the precise model inferred for later simulation
is at best tedious and error prone~\citep{adrion2020community,ragsdale2020lessons},
and occasionally impossible because of missing information.
% \mhcomment{Maybe add a short, concrete example to convince readers of the problem?}
% APR: I think this is ok here.
% Instead, it would be far better if published results were reported
% in an unambiguous standardised format that could be provided
% directly as input to simulators.

Simulation is a core tool in population genetics, and
many methods have been developed over the past three
decades~\citep{carvajal2008simulation,liu2008survey,
arenas2012simulation,yuan2012overview,hoban2012computer}.
Simulations are based on highly idealised population models,
and one of the key uses of inferred demographic histories
is to make simulations more realistic.
Simulation methods take three broad approaches to specifying
the demographic model to simulate,
using either a command line
interface~\citep[e.g.,][]{hudson2002generating,hernandez_flexible_2008,kern2016discoal},
a custom input file
format~\citep[e.g.,][]{guillaume2006nemo,excoffier2011fastsimcoal,shlyakhter2014cosi2},
or an Application Programming Interface (API) to allow
models to be defined programmatically~\citep[e.g.,][]{
thornton2014cpp,hernandez_sfs_code_2015,kelleher2016efficient,
becheler2019quetzal,haller2019slim,thornton2019-nu,baumdicker2021-iu}.
Command line interfaces are a concise way of expressing
demographic models, and the syntax defined by \ms~\citep{hudson2002generating}
is used by a number of different
simulators~\citep[e.g.,][]{ewing2010msms,chen2009fast,staab2015scrm}.
However, this conciseness means that models of even intermediate complexity
are difficult for humans to understand, making errors likely.
Input parameter file formats for simulators allow the model specification
to be less terse and allow for documentation in the form of comments.
Several Graphical User Interface (GUI) and visualisation
methods have been developed, which greatly facilitate
interpretation~\citep{mailund2005coasim,antao2007modeler4simcoal2,
parreira_spams_2009,ewing2010msms,parobek_skelesim_2017,zhou2018popdemog}.
However, these methods
currently have little traction as they are all either directly coupled
to an internal simulation method or to the  syntax of a specific simulator.
There is currently no way in which demographic models inferred by
different packages can be can be simulated or visualised by downstream software.

@@

Population sizes are given as numbers of individuals, and details
such as ploidy levels are considered external to the model.
We therefore focus on the number of individuals as opposed
to the number of genome copies.
Sizes and mating system details are specified for each population within
population-specific epochs.
Epochs are contiguous time intervals that define
the existence interval of the population. Each epoch specifies the population size
over that interval, which can be a constant value or function defined by start
and end sizes that must remain positive.

>>>>

% local definitions
\newcommand{\ms}[0]{\texttt{ms}}
\newcommand{\msprime}[0]{\texttt{msprime}}
\newcommand{\stdpopsim}[0]{\texttt{stdpopsim}}
\newcommand{\demes}[0]{\texttt{demes}}
\newcommand{\demesdraw}[0]{\texttt{demesdraw}}
\newcommand{\Demes}[0]{\texttt{Demes}}
\newcommand{\moments}[0]{\texttt{moments}}
\newcommand{\dadi}[0]{\texttt{dadi}}
\newcommand{\fwdpy}[0]{\texttt{fwdpy11}}
\newcommand{\slim}[0]{\texttt{SLiM}}
\newcommand{\gadma}[0]{\texttt{GADMA}}
\newcommand{\tskit}[0]{\texttt{tskit}}
\newcommand{\demesslim}[0]{\texttt{demes-slim}}

@@

The ever-increasing amount of genetic sequencing data from genetically and
geographically diverse species and populations has allowed us to infer complex
demography and study life history at fine scales.
An integral component to such population genetics studies is simulation.
Software to either simulate whole genome
sequences~\citep{thornton2014cpp,thornton2019-nu,staab2015scrm,
baumdicker2021-iu,kelleher2016efficient,haller2019slim}
or informative summary statistics of
diversity~\citep{gutenkunst2009inferring,kamm2017efficient,jouganous2017inferring}
have enabled the increasing complexity of genomic studies, with several software
packages capable of handling large sample sizes, many interacting populations, and
deviations from panmictic random-mating assumptions.
This ability to infer and simulate such complex demographic scenarios, however,
has highlighted a major shortcoming in community standards:
the fragmented landscape of different ways to describe demographic
models makes it difficult to compare inferences made by different methods
and to reliably simulate from previously inferred models.
Inference results are typically reported in publications
via a combination of visual depiction,
a list of key parameters in tabular form and a discussion within the text.
Unfortunately these descriptions are often ambiguous, and
implementing the precise model inferred for later simulation
is at best tedious and error prone~\citep{adrion2020community,ragsdale2020lessons},
and occasionally impossible because of missing information.
% \mhcomment{Maybe add a short, concrete example to convince readers of the problem?}
% APR: I think this is ok here.
% Instead, it would be far better if published results were reported
% in an unambiguous standardised format that could be provided
% directly as input to simulators.

Simulation is a core tool in population genetics, and
many methods have been developed over the past three
decades~\citep{carvajal2008simulation,liu2008survey,
arenas2012simulation,yuan2012overview,hoban2012computer}.
Simulations are based on highly idealised population models,
and one of the key uses of inferred demographic histories
is to make simulations more realistic.
Simulation methods take three broad approaches to specifying
the demographic model to simulate,
using either a command line
interface~\citep[e.g.,][]{hudson2002generating,hernandez_flexible_2008,kern2016discoal},
a custom input file
format~\citep[e.g.,][]{guillaume2006nemo,excoffier2011fastsimcoal,shlyakhter2014cosi2},
or an Application Programming Interface (API) to allow
models to be defined programmatically~\citep[e.g.,][]{
thornton2014cpp,hernandez_sfs_code_2015,kelleher2016efficient,
becheler2019quetzal,haller2019slim,thornton2019-nu,baumdicker2021-iu}.
Command line interfaces are a concise way of expressing
demographic models, and the syntax defined by \ms~\citep{hudson2002generating}
is used by a number of different
simulators~\citep[e.g.,][]{ewing2010msms,chen2009fast,staab2015scrm}.
However, this conciseness means that models of even intermediate complexity
are difficult for humans to understand, making errors likely.
Input parameter file formats for simulators allow the model specification
to be less terse and allow for documentation in the form of comments.
Several Graphical User Interface (GUI) and visualisation
methods have been developed, which greatly facilitate
interpretation~\citep{mailund2005coasim,antao2007modeler4simcoal2,
parreira_spams_2009,ewing2010msms,parobek_skelesim_2017,zhou2018popdemog}.
However, these methods
currently have little traction as they are all either directly coupled
to an internal simulation method or to the  syntax of a specific simulator.
There is currently no way in which demographic models inferred by
different packages can be simulated or visualised by downstream software.

@@

Population sizes are given as numbers of individuals, and details
such as ploidy levels are considered external to the model.
We therefore focus on the number of individuals as opposed
to the number of genome copies.
Sizes and mating system details are specified for each population within
population-specific epochs.
Epochs are contiguous time intervals that define
the existence interval of the population. Each epoch specifies the population size
over that interval, which can be a constant value or a function defined by start
and end sizes that must remain positive.
